---
title: "Technical appendix: Networked Data Lab analysis on unpaid carers"
author: "The Health Foundation"
date: "`r Sys.Date()`"
output:
  rmarkdown::html_document:
    theme: lumen
    toc: true
---



```{r, echo=FALSE, message=FALSE, warning=FALSE}
#Load packages

library(aws.s3)
library(tidyverse)
library(data.table)
library(kableExtra)
library(knitr)

#Clean up the global environment

rm(list = ls())

#Directories in S3

IHT_bucket <- "s3://thf-dap-tier0-projects-iht-067208b7-projectbucket-1mrmynh0q7ljp"
ASC_subfolder <- "ASC and Finance Report"
R_workbench <- path.expand("~")
localgit <- dirname(rstudioapi::getSourceEditorContext()$path)

clinicalcodes <- s3read_using(fread,
                                          object = "NDL-carers-partner-data/Consolidated codes.csv",
                                          bucket = IHT_bucket)
```

## The Networked Data Lab

The Networked Data Lab (NDL) is a collaborative network of five analytical teams across the UK working together using linked datasets on key issues facing health and care services today. This analysis was on support services delivered to unpaid carers, and involved working with health and social care data from the following local authorities:

- Neath Port Talbot (Wales)
- Swansea (Wales)
- Liverpool (England)
- Wirral (England)
- Leeds (England)

The following local authorities only provided GP data, and are known collectively as North West London (England):

- Brent
- Westminster
- Ealing
- Hammersmith and Fulham
- Harrow
- Hillingdon
- Hounslow
- West London

## Federated analysis approach

The results presented in our briefing use a federated analytics model whereby analyses were carried out by each of the Networked Data Lab partners without any patient-level data leaving any of their research environments. While all partners researched similar topics, each partner designed and followed their own analysis plan, all of which are detailed here. 

## Inclusion criteria

- We included all unpaid carers over the **age of 18** who had either a GP or adult social care record during the study period, indicating they were an unpaid carers.

- **GPs** (or GP practice staff) use clinical codes to keep a record of their patients who are unpaid carers. This is so they can support unpaid carers in taking care of their own health and the health of the person (or people) they are caring for. A limitation of clinical codes is that they have an unknown half-life: they show when a GP finds out someone is a carer, but not when they started being a carer and they don't reveal when someone is no longer a carer.

- Similarly, **local authorities** with **adult social care** responsibilities keep records of unpaid carers who come forward to request support - for example, to provide respite care or direct payments. Carers must first undergo a carer's assessment, which determines whether they are eligible for these types of support. Each contact is associated with a date (for example, when they first reach out to request support or when support was delivered). It is likely that people have been carers for some time before becoming known to adult social care services. Similarly, these datasets do not reveal when someone is no longer a carer.

- We started by removing all records before the start date or after the end date. We also remove all records associated with a person under the age of 18 at the time of the record.

## Handling of index dates

### Pooled analysis (all years pooled together)

- Each person is associated with a single **index date**, which is the date of their earliest valid record found (GP or adult social care flag indicating they are an unpaid carer).  

### Yearly analysis (one cohort per year)

- If a person has records in multiple years (for example, 2 GP codes in 2018 and 1 GP code in 2019) then they are included in the cohorts for multiple years (for example, 2018 and 2019 cohorts). The **index date** for each yearly cohort is the first valid record within that year.

## Study dates

Due to differences in data availability, not all local authorities have the same study start and end dates.

- Swansea: 01/04/2021 to 19/06/2022 (1 year)
- Neath Port Talbot: 01/07/2017 to 26/06/2022 (5 years)
- Liverpool: 01/01/2016 to 31/12/2021 (6 years)
- Wirral: 01/01/2016 to 31/12/2021 (6 years)
- Leeds: 01/01/2016 to 31/12/2021 (6 years)
- North West London: 01/01/2016 to 31/12/2021 (6 years)

## Clinical code lists

Here below are the clinical code lists used by each of the Networked Data Lab partners to identify unpaid carers in GP data.

```{r, eval=TRUE, echo=FALSE, message=FALSE, warning=FALSE}
clinicalcodes %>%
  kbl() %>%
  kable_styling(bootstrap_options = "striped", full_width = F, position = "left")
```